---
title: "Ecoregion (State Child)"
author: "Rowan R"
date: "2025-12-07"
output: 
  pdf_document:
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
### functions ----
perPlnt_df <- function(df){
  perPlnt <- df %>% 
   # filter(name %in% ecor) %>% 
    group_by(name, plantGenSpp, orig1) %>% 
    summarise(nBees = sum(n),
              nBeeSpp = n_distinct(genSpp), 
              beeSpp = paste0(genSpp, collapse = ", "))
  return(perPlnt) 
}

top10 <- function(df, varName){
  dO <- df %>% 
    arrange(desc({{varName}})) %>% 
    head(n = 10)
  return(dO)
}
```

```{r, include=FALSE, warning=FALSE}

singleEco <- datEcoR %>% 
  filter(name %in% x) 

singleEcoNoMissing <- singleEco %>% 
  st_drop_geometry() %>% 
  filter(!is.na(genSpp)) %>% 
  filter(!is.na(plantGenSpp)) %>% 
  mutate(monthNum = as.numeric(month), 
         dayNum = as.numeric(day))

splitSeasons <- function(df, seas = num1to3){
  if(seas %in% 1){ # SPRING - winter currently excluded
    dfO <- df %>% filter(between(monthNum, left = 4, right = 5)|(monthNum %in% 3 & dayNum>20)|(monthNum %in% 6 & dayNum<21) )
  }else if(seas %in% 2) { # SUMMER
      dfO <- df %>% filter(between(x = monthNum, left = 7, right = 8)|(monthNum %in% 6 & dayNum>20)|monthNum %in% 9 & dayNum < 23)
  }else if(seas %in% 3){ # FALL
      dfO <- df %>% filter(between(monthNum, 10, 11)|(monthNum %in% 9 & dayNum > 22)|(monthNum %in% 12 & dayNum < 22))
    }else print("Not a recognized season, options are 1, 2, 3")
  return(dfO)
}

ecot1 <- splitSeasons(singleEcoNoMissing, seas = 1)
ecot2 <-  splitSeasons(singleEcoNoMissing, seas = 2)
ecot3 <-  splitSeasons(singleEcoNoMissing, seas = 3)

# Only use this data for doing interactions, not species overall etc. since males are removed
singleEcoInt <- singleEcoNoMissing %>% 
  filter(!is.na(genSpp)) %>% 
  filter(!is.na(plantGenSpp)) %>% 
  filter(!sex %in% c("Male", "male")) %>% # removing males for now - not informative for pollen
  group_by(name, genSpp, plantGenSpp, orig1) %>% # might adjust this later - using plant genus
  count()

singleEcoPerPlnt <- perPlnt_df(singleEcoInt)

singleEcoPerPlnt1 <- perPlnt_df(
 ecot1 %>% 
  filter(!is.na(genSpp)) %>% 
  filter(!is.na(plantGenSpp)) %>% 
  filter(!sex %in% c("Male", "male")) %>% # removing males for now - not informative for pollen
  group_by(name, genSpp, plantGenSpp, orig1) %>% # might adjust this later - using plant genus
  count()
)

singleEcoPerPlnt2 <- perPlnt_df(
 ecot2 %>% 
  filter(!is.na(genSpp)) %>% 
  filter(!is.na(plantGenSpp)) %>% 
  filter(!sex %in% c("Male", "male")) %>% # removing males for now - not informative for pollen
  group_by(name, genSpp, plantGenSpp, orig1) %>% # might adjust this later - using plant genus
  count()
)

singleEcoPerPlnt3 <- perPlnt_df(
 ecot3 %>% 
  filter(!is.na(genSpp)) %>% 
  filter(!is.na(plantGenSpp)) %>% 
  filter(!sex %in% c("Male", "male")) %>% # removing males for now - not informative for pollen
  group_by(name, genSpp, plantGenSpp, orig1) %>% # might adjust this later - using plant genus
  count()
)

```

\newpage
\section {Ecoregion: `r x`}
```{r, include=FALSE, warning=FALSE}
nSppB_e <- n_distinct(singleEco$genSpp)

nSppP_e <- n_distinct(singleEcoNoMissing$plantGenSpp)

sppInts <- singleEcoNoMissing %>% mutate(ints = str_c(genSpp, "-", plantGenSpp))
nInts_e <- n_distinct(sppInts$ints)

```

Total Bee Species: `r nSppB_e`

Total Plant species: `r nSppP_e`

Total Bee-Plant Interactions: `r nInts_e`

```{r, include=FALSE}
# side by side per county? with county highlighted
ecor2highlight <- stateChao_ecoreg %>% filter(name %in% x) %>% 
  mutate(gap = as.character(round(Estimator - Observed), 0))

# make figure caption
fCap <- str_c("Choropleth of documented bee species richness for level III ecoregions in ", stateName, " with the ", x, " ecoregion outlined. Dots on the map represent the locations of all sampling localities within the ", x, " ecoregion.")
```


```{r, include = FALSE} 
ecoQuad <- st_centroid(stateChao_ecoreg) %>% st_join(quadrants, st_within) %>% print(n = 40) %>% 
  select(name, NA_L3NAME, quad) %>% st_drop_geometry()

stateChao_ecoreg <- stateChao_ecoreg %>% left_join(ecoQuad)
whichEcoQuad <- ecor2highlight %>% left_join(ecoQuad) %>% pull(quad)
ecorSubset <- stateChao_ecoreg %>% ungroup() %>% st_filter(., ecor2highlight, .predicate = st_intersects)
```

\subsection {Ecoregion Summary Map}

```{r, fig.cap= fCap, fig.asp= 0.75}
ecoSppPlot <- actualStateRich(stateChao_ecoreg)+
  scale_fill_gradient2(low = "white",
                       mid = '#f1eef6',
                       high = '#045a8d')+
  geom_sf(data = ecor2highlight, colour = "green", alpha = 0, size = 1)+
    labs(x = "", y = "")+
  geom_sf(data = datEcoR %>% filter(name %in% x), size = 0.75)+
  geom_sf_text(data = ecor2highlight, aes(label = Observed), size = 4, color = "yellow")+
  labs(fill = "Documented \n   Species")

cowplot::plot_grid(ecoSppPlot)

# cowplot::plot_grid(
# actualStateRich(stateChao_ecoreg)+
#   scale_fill_gradient2(low = "white", 
#                        mid = '#f1eef6',
#                        high = '#045a8d')+
#   geom_sf(data = ecor2highlight, colour = "green", alpha = 0, size = 1)+
#     geom_sf_text(data = ecor2highlight, aes(label = Observed), size = 2)+
#     labs(x = "", y = "") , 
# gapStateRich(stateChao_ecoreg)+
#   scale_fill_gradient2(low = "white", 
#                        mid = '#ffffcc',
#                        high = '#bd0026')+
#   geom_sf(data = ecor2highlight, colour = "green", alpha = 0, size = 1)+
#   geom_sf_text(data = ecor2highlight, aes(label = gap), size = 2)+
#   labs(x = "", y = ""),
# 
# nrow = 2
# )
```


\newpage


```{r, include=FALSE}
tCap <- str_c("Top sampled plants for documented bee richness in the ", x, " ecoregion.") 
```

\subsection{Top Plants For Bee Richness}


```{r, echo=FALSE, fig.cap= tCap, fig.width=6.5, fig.height= 5.5}

tpgen <- slice_head(singleEcoPerPlnt %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>% 
            rename(`Plant Species` = plantGenSpp, Origin = orig1, `Documented Bee Visits` = nBees,`Bee Species` = nBeeSpp), n = 10)

ggplot(tpgen %>% rename(`Bee Species Documented` = `Bee Species`) %>% pivot_longer(cols = contains("Bee"), names_to = "Number", values_to = "val"), aes(x = `Plant Species`, y = `val`, fill = Origin))+
  geom_col()+
  facet_wrap(~Number, scales = "free_x")+
  coord_flip()+
  scale_fill_brewer(palette = "Paired")+
  geom_text(aes(label = `val`), position = position_stack(vjust = 0.5))+
  theme(axis.text.y = element_text(face = "italic"))+ 
  labs(y = "Number")

```


\newpage

\subsection {Top plants for bee richness, split by season}

```{r, include=FALSE}

tpgen1 <- slice_head(singleEcoPerPlnt1 %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>% 
            rename(`Plant Species` = plantGenSpp, Origin = orig1, `Number Bee Visits Observed` = nBees,`Bee Species` = nBeeSpp) %>% 
  mutate(seas = "Spring"), n = 10)

tpgen2 <- slice_head(singleEcoPerPlnt2 %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>% 
            rename(`Plant Species` = plantGenSpp, Origin = orig1, `Number Bee Visits Observed` = nBees,`Bee Species` = nBeeSpp) %>% 
              mutate(seas = "Summer"),  n = 10)
  
tpgen3 <- slice_head(singleEcoPerPlnt3 %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>% 
            rename(`Plant Species` = plantGenSpp, Origin = orig1, `Number Bee Visits Observed` = nBees,`Bee Species` = nBeeSpp) %>% 
              mutate(seas = "Fall"),  n = 10)

tpgenSeas <- rbind(tpgen1, tpgen2, tpgen3) %>% mutate(`Number Bee Visits Observed` = as.integer(`Number Bee Visits Observed`), `Bee Species` = as.integer(`Bee Species`))
```

```{r, fig.height=7.5, fig.width = 10, fig.cap = str_c("Top sampled plant species that support the most documented bee species by season in the ", x, " ecoregion. Seasons are defined as spring (Mar 21 - Jun 20), summer (Jun 21 - Sep 22), and fall (Sep 23 - Dec 21).") }
ggplot(tpgenSeas %>% rename(`Bee Spp. Doc.` = `Bee Species`, `No. Bee Visits` = `Number Bee Visits Observed`) %>% pivot_longer(cols = contains("Bee"), names_to = "Number", values_to = "val") %>% 
         mutate(seas = fct_relevel(as_factor(seas), "Spring", "Summer", "Fall")), aes(x = `Plant Species`, y = `val`, fill = Origin))+
  geom_col()+
  facet_wrap(~Number*seas, scales = "free")+
  coord_flip()+
  scale_fill_brewer(palette = "Paired")+
  geom_text(aes(label = `val`), position = position_stack(vjust = 0.5))+
  theme(axis.text.y = element_text(face = "italic"),
        panel.spacing.y = unit(2, "lines"))+ 
  labs(y = "Number")+
  scale_y_continuous(labels = scales::label_number(accuracy = 1), n.breaks = 5)

```

<!-- ```{r, include=FALSE} -->
<!-- tcap_1 <- paste0("Top plant genera for bee richness in the ", x, " Ecoregion during winter and early spring (Jan - March).") -->
<!-- tcap_2 <- paste0("Top plant genera for bee richness in the ", x, " Ecoregion during spring (April - June).") -->
<!-- tcap_3 <- paste0("Top plant genera for bee richness in the ", x, " Ecoregion during Summer (July - October).") -->
<!-- ``` -->

<!-- ```{r, tab.cap= tcap_1} -->
<!-- kable(slice_head(singleEcoPerPlnt1 %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>%  -->
<!--                   rename(`Plant Species` = plantGenSpp, Origin = orig1, `Number Bee Visits Observed` = nBees,`Bee Species` = nBeeSpp),  n = 10)) -->
<!-- ``` -->

<!-- ```{r, tab.cap= tcap_2} -->
<!-- kable(slice_head(singleEcoPerPlnt2 %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>%  -->
<!--                   rename(`Plant Species` = plantGenSpp, Origin = orig1, `Number Bee Visits Observed` = nBees,`Bee Species` = nBeeSpp),  n = 10)) -->

<!-- ``` -->

<!-- ```{r, tab.cap= tcap_3} -->
<!-- kable(slice_head(singleEcoPerPlnt3 %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>%  -->
<!--                   rename(`Plant Species` = plantGenSpp, Origin = orig1, `Number Bee Visits Observed` = nBees,`Bee Species` = nBeeSpp),  n = 10)) -->

<!-- ``` -->
\newpage
\subsection {Most abundant bee species of the `r x` ecoregion}

```{r, include=FALSE}
tcap_topb_1 <- str_c("Most abundant bee species documented in the ", x, " ecoregion by season. Seasons are defined as spring (Mar 21 - Jun 20), summer (Jun 21 - Sep 22), and fall (Sep 23 - Dec 21).")
#tcap_topb_2 <- paste0("Most Abundant Bee Species of the ", x, " ecoregion during spring (April - June).")
#tcap_topb_3 <- paste0("Most Abundant Bee Species of the ", x, " ecoregion during Summer (July - October).")
```

```{r, fig.cap= tcap_topb_1, fig.width= 9.5, fig.height = 4}
abun1 <- slice_head(ecot1 %>% filter(!str_detect(genSpp, "spp\\.")) %>% group_by(genSpp, family) %>% count() %>% 
                   ungroup() %>% arrange(desc(n)) %>% rename(Species = genSpp, Observations = n) %>% 
                     mutate(seas = "Spring"), n = 10)

abun2 <- slice_head(ecot2 %>% filter(!str_detect(genSpp, "spp\\.")) %>% group_by(genSpp, family) %>% count() %>% 
                   ungroup() %>% arrange(desc(n)) %>% rename(Species = genSpp, Observations = n) %>% 
                     mutate(seas = "Summer"), n = 10) 

abun3 <- slice_head(ecot3 %>% filter(!str_detect(genSpp, "spp\\.")) %>% group_by(genSpp, family) %>% count() %>% 
                   ungroup() %>% arrange(desc(n)) %>% rename(Species = genSpp, Observations = n) %>% 
                     mutate(seas = "Fall"),          n = 10)

abunAll <- rbind(abun1, abun2, abun3) %>% 
  mutate(seas = fct_relevel(as_factor(seas), "Spring", "Summer", "Fall"))

famColSimple_df <- tibble(cols=c(brewer.pal(5,'Set1'), "black"), 
                  family = c('Andrenidae','Apidae','Colletidae',
                           'Halictidae','Megachilidae', "Melittidae"))

famColSimple <- as.character(famColSimple_df$cols)
names(famColSimple) <- famColSimple_df$family

ggplot(abunAll, aes(x = Species, y = Observations, fill = family ))+
  geom_col()+
  scale_fill_manual(values = famColSimple)+
  coord_flip()+
  theme(axis.text.y = element_text(face = "italic"))+
  labs(fill = "Bee Family")+
  facet_wrap(~seas, scales = "free")+
  scale_y_continuous(labels = scales::label_number(accuracy = 1), n.breaks = 4)
```

<!-- ```{r, tab.cap= tcap_topb_1} -->
<!-- kable(slice_head(ecot1 %>% filter(!str_detect(genSpp, "spp\\.")) %>% group_by(genSpp) %>% count() %>%  -->
<!--                    ungroup() %>% arrange(desc(n)) %>% rename(Species = genSpp, Observations = n),  -->
<!--                  n = 10) ) # REPLACED WITH ABOVE FIGURES -->
<!-- ``` -->

<!-- ```{r, tab.cap= tcap_topb_2} -->
<!-- kable(slice_head(ecot2 %>% filter(!str_detect(genSpp, "spp\\.")) %>% group_by(genSpp) %>% count() %>%  -->
<!--                    ungroup() %>% arrange(desc(n)) %>% rename(Species = genSpp, Observations = n),  -->
<!--                  n = 10) ) -->
<!-- ``` -->

<!-- ```{r, tab.cap= tcap_topb_3} -->
<!-- kable(slice_head(ecot3 %>% filter(!str_detect(genSpp, "spp\\.")) %>% group_by(genSpp) %>% count() %>%  -->
<!--                    ungroup() %>% arrange(desc(n)) %>% rename(Species = genSpp, Observations = n),  -->
<!--                  n = 10) ) -->
<!-- ``` -->

\subsection {Uncommon Bee Species for the `r x` ecoregion}
```{r, include=FALSE}
tcap_unc <- paste0("Bees with fewer than 15 ", stateName, " Bee Atlas observations, and fewer than 10 in the ",  x, " ecoregion.")
```


```{r, tab.cap= tcap_unc}
uncft <- flextable(singleEcoNoMissing %>% 
  group_by(genSpp) %>% 
  count() %>% ungroup() %>% 
  semi_join(below15_proj, by = "genSpp") %>% 
    rename(`Bee Species` = genSpp, Observations = n)
  ) %>% 
  italic(j = 1) %>% 
  fontsize(size = 10) %>% 
  align( j = 2, align = "center", part = "all") %>% 
  autofit()

# uncftwids <- dim(uncft)$widths/sum(dim(uncft)$widths)*6.5
# 
# uncft <- width(uncft, width = uncftwids)
uncft
```
\subsection {Known specialist bee species present in the `r x` ecoregion}

 This list is produced by querying species included in the literature review of specialist bees produced by Fowler and Droege (2020). As this list is derived from historical literature, there are errors both in underestimating and overestimating specialization, and some misapplications of plant names. Many more specialist bees have been documented by the Oregon Bee Atlas, but these are omitted because their floral relations have not yet been recognized in the literature, or were overlooked in the production of our reference list. Empirical analysis of Oregon Bee Atlas bee-plant interaction occurrence records will add many new specialized relationships to this output in future publications.
 
```{r, tab.cap= str_c("Specialist bees present in the ", x, " ecoregion with the number of times each species has been documented and its reported floral host(s).")}
library(flextable)

specft <- flextable(singleEcoNoMissing %>% group_by(genSpp) %>% count() %>% ungroup() %>%  
         semi_join(specBeesLookup) %>% left_join(specBeesLookup %>% select(genSpp, `Host plant Family Tribe Genera`)) %>% 
  rename(`Bee Species` = genSpp, Obs. = n, `Floral Hosts (Fowler & Droege, 2020)` = `Host plant Family Tribe Genera` )) 

specft  <- autofit(specft)
auto_wid <- dim(specft)$widths/sum(dim(specft)$widths)
auto_wid[1] <- 1.5
auto_wid[2] <- 0.5
#auto_wid[3] <- 1
auto_wid[3] <- 3.5 # specify widths - auto isn't working for this case

specft <- width(specft, width = auto_wid)

library(officer)
inBorder <- fp_border(color = "lightgrey", width = 0.5)

specft %>% 
  italic( j = 1) %>% 
  align(j = 2, align = "center", part = "all") %>% 
  fontsize(size = 9, part = "body") %>% 
  fontsize(size = 10, part = "header") %>% 
  border_inner_h(border = inBorder, part = "body") 
```

