---
title: "Ecoregion (State Child)"
author: "Rowan R"
date: "2025-12-07"
output: 
  pdf_document:
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
### functions ----
perPlnt_df <- function(df){
  perPlnt <- df %>% 
   # filter(name %in% ecor) %>% 
    group_by(name, plantGenSpp, orig1) %>% 
    summarise(nBees = sum(n),
              nBeeSpp = n_distinct(genSpp), 
              beeSpp = paste0(genSpp, collapse = ", "))
  return(perPlnt) 
}

top10 <- function(df, varName){
  dO <- df %>% 
    arrange(desc({{varName}})) %>% 
    head(n = 10)
  return(dO)
}
```

```{r, include=FALSE, warning=FALSE}

singleEco <- datEcoR %>% 
  filter(name %in% x) 

singleEcoNoMissing <- singleEco %>% 
  st_drop_geometry() %>% 
  filter(!is.na(genSpp)) %>% 
  filter(!is.na(plantGenSpp)) %>% 
  mutate(monthNum = as.numeric(month))

splitSeasons <- function(df, earlyMid = 4, midLate = 7, seas = num1to3){
  if(seas %in% 1){
    dfO <- df %>% filter(monthNum < earlyMid)
  }else if(seas %in% 2) {
      dfO <- df %>% filter(between(x = monthNum, left = earlyMid, right = midLate))
  }else if(seas %in% 3){
      dfO <- df %>% filter(monthNum > midLate)
    }else print("Not a recognized season, options are 1, 2, 3")
  return(dfO)
}

ecot1 <- splitSeasons(singleEcoNoMissing, seas = 1)
ecot2 <-  splitSeasons(singleEcoNoMissing, seas = 2)
ecot3 <-  splitSeasons(singleEcoNoMissing, seas = 3)

# Only use this data for doing interactions, not species overall etc. since males are removed
singleEcoInt <- singleEcoNoMissing %>% 
  filter(!is.na(genSpp)) %>% 
  filter(!is.na(plantGenSpp)) %>% 
  filter(!sex %in% c("Male", "male")) %>% # removing males for now - not informative for pollen
  group_by(name, genSpp, plantGenSpp, orig1) %>% # might adjust this later - using plant genus
  count()

singleEcoPerPlnt <- perPlnt_df(singleEcoInt)

singleEcoPerPlnt1 <- perPlnt_df(
 ecot1 %>% 
  filter(!is.na(genSpp)) %>% 
  filter(!is.na(plantGenSpp)) %>% 
  filter(!sex %in% c("Male", "male")) %>% # removing males for now - not informative for pollen
  group_by(name, genSpp, plantGenSpp, orig1) %>% # might adjust this later - using plant genus
  count()
)

singleEcoPerPlnt2 <- perPlnt_df(
 ecot2 %>% 
  filter(!is.na(genSpp)) %>% 
  filter(!is.na(plantGenSpp)) %>% 
  filter(!sex %in% c("Male", "male")) %>% # removing males for now - not informative for pollen
  group_by(name, genSpp, plantGenSpp, orig1) %>% # might adjust this later - using plant genus
  count()
)

singleEcoPerPlnt3 <- perPlnt_df(
 ecot3 %>% 
  filter(!is.na(genSpp)) %>% 
  filter(!is.na(plantGenSpp)) %>% 
  filter(!sex %in% c("Male", "male")) %>% # removing males for now - not informative for pollen
  group_by(name, genSpp, plantGenSpp, orig1) %>% # might adjust this later - using plant genus
  count()
)

```

\newpage
\section {Ecoregion: `r x`}
```{r, include=FALSE, warning=FALSE}
nSppB_e <- n_distinct(singleEco$genSpp)

nSppP_e <- n_distinct(singleEcoNoMissing$plantGenSpp)

sppInts <- singleEcoNoMissing %>% mutate(ints = str_c(genSpp, "-", plantGenSpp))
nInts_e <- n_distinct(sppInts$ints)

```

Total Bee Species: `r nSppB_e`

Total Plant species: `r nSppP_e`

Total Bee-Plant Interactions: `r nInts_e`

```{r, include=FALSE}
tCap <- str_c("Top plant species/genera for bee richness in the ", x, " Ecoregion.") 
```

\subsection {Top Plant Genera For Bee Richness}

```{r, echo=FALSE, tab.cap= str_c("To replace with below figure? ", tCap)}
tpgen <- slice_head(singleEcoPerPlnt %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>% 
                  rename(`Plant Species` = plantGenSpp, Origin = orig1, `Number Bee Visits Observed` = nBees,`Bee Species` = nBeeSpp),  n = 10)

flextable(tpgen) %>% 
  italic(j = 1) %>%
  align(part = "all", j = 2:4, align = "center")
```
\newpage

```{r, echo=FALSE, fig.cap= tCap}

ggplot(tpgen %>% rename(`Bee Species Observed` = `Bee Species`) %>% pivot_longer(cols = contains("Bee"), names_to = "Number", values_to = "val"), aes(x = `Plant Species`, y = `val`, fill = Origin))+
  geom_col()+
  facet_wrap(~Number, scales = "free_x")+
  coord_flip()+
  scale_fill_brewer(palette = "Paired")+
  geom_text(aes(label = `val`), position = position_stack(vjust = 0.5))+
  theme(axis.text.y = element_text(face = "italic"))+ 
  labs(y = "Number")

```
\newpage

\subsection {Top plant genera for bee richness, split by season}

```{r, include=FALSE}
tpgen1 <- slice_head(singleEcoPerPlnt1 %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>% 
            rename(`Plant Species` = plantGenSpp, Origin = orig1, `Number Bee Visits Observed` = nBees,`Bee Species` = nBeeSpp) %>% 
  mutate(seas = "Jan - March"), n = 10)

tpgen2 <- slice_head(singleEcoPerPlnt2 %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>% 
            rename(`Plant Species` = plantGenSpp, Origin = orig1, `Number Bee Visits Observed` = nBees,`Bee Species` = nBeeSpp) %>% 
              mutate(seas = "April - June"),  n = 10)
  
tpgen3 <- slice_head(singleEcoPerPlnt3 %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>% 
            rename(`Plant Species` = plantGenSpp, Origin = orig1, `Number Bee Visits Observed` = nBees,`Bee Species` = nBeeSpp) %>% 
              mutate(seas = "July - October"),  n = 10)

tpgenSeas <- rbind(tpgen1, tpgen2, tpgen3)
```

```{r, fig.height=6.5, fig.width = 10, fig.cap = str_c("Plant species that support the most bee species by season in the ", x, " Ecoregion.") }
ggplot(tpgenSeas %>% rename(`Bee Spp. Detected` = `Bee Species`, `No. Bee Visits` = `Number Bee Visits Observed`) %>% pivot_longer(cols = contains("Bee"), names_to = "Number", values_to = "val") %>% 
         mutate(seas = fct_relevel(as_factor(seas), "Jan - March", "April - June", "July - October")), aes(x = `Plant Species`, y = `val`, fill = Origin))+
  geom_col()+
  facet_wrap(~Number*seas, scales = "free")+
  coord_flip()+
  scale_fill_brewer(palette = "Paired")+
  geom_text(aes(label = `val`), position = position_stack(vjust = 0.5))+
  theme(axis.text.y = element_text(face = "italic"))+ 
  labs(y = "Number")

```

<!-- ```{r, include=FALSE} -->
<!-- tcap_1 <- paste0("Top plant genera for bee richness in the ", x, " Ecoregion during winter and early spring (Jan - March).") -->
<!-- tcap_2 <- paste0("Top plant genera for bee richness in the ", x, " Ecoregion during spring (April - June).") -->
<!-- tcap_3 <- paste0("Top plant genera for bee richness in the ", x, " Ecoregion during Summer (July - October).") -->
<!-- ``` -->

<!-- ```{r, tab.cap= tcap_1} -->
<!-- kable(slice_head(singleEcoPerPlnt1 %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>%  -->
<!--                   rename(`Plant Species` = plantGenSpp, Origin = orig1, `Number Bee Visits Observed` = nBees,`Bee Species` = nBeeSpp),  n = 10)) -->
<!-- ``` -->

<!-- ```{r, tab.cap= tcap_2} -->
<!-- kable(slice_head(singleEcoPerPlnt2 %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>%  -->
<!--                   rename(`Plant Species` = plantGenSpp, Origin = orig1, `Number Bee Visits Observed` = nBees,`Bee Species` = nBeeSpp),  n = 10)) -->

<!-- ``` -->

<!-- ```{r, tab.cap= tcap_3} -->
<!-- kable(slice_head(singleEcoPerPlnt3 %>% arrange(desc(nBeeSpp)) %>%ungroup() %>% select(!c(name, beeSpp)) %>%  -->
<!--                   rename(`Plant Species` = plantGenSpp, Origin = orig1, `Number Bee Visits Observed` = nBees,`Bee Species` = nBeeSpp),  n = 10)) -->

<!-- ``` -->
\newpage
\subsection {Most abundant bee species of the `r x` Ecoregion}

```{r, include=FALSE}
tcap_topb_1 <- str_c("Most abundant bee species of the ", x, " Ecoregion by season.")
#tcap_topb_2 <- paste0("Most Abundant Bee Species of the ", x, " Ecoregion during spring (April - June).")
#tcap_topb_3 <- paste0("Most Abundant Bee Species of the ", x, " Ecoregion during Summer (July - October).")
```

```{r, fig.cap= tcap_topb_1, fig.width= 9.5, fig.height = 4}
abun1 <- slice_head(ecot1 %>% filter(!str_detect(genSpp, "spp\\.")) %>% group_by(genSpp, family) %>% count() %>% 
                   ungroup() %>% arrange(desc(n)) %>% rename(Species = genSpp, Observations = n) %>% 
                     mutate(seas = "Jan - March"), n = 10)

abun2 <- slice_head(ecot2 %>% filter(!str_detect(genSpp, "spp\\.")) %>% group_by(genSpp, family) %>% count() %>% 
                   ungroup() %>% arrange(desc(n)) %>% rename(Species = genSpp, Observations = n) %>% 
                     mutate(seas = "April - June"), n = 10) 

abun3 <- slice_head(ecot3 %>% filter(!str_detect(genSpp, "spp\\.")) %>% group_by(genSpp, family) %>% count() %>% 
                   ungroup() %>% arrange(desc(n)) %>% rename(Species = genSpp, Observations = n) %>% 
                     mutate(seas = "July - October"),          n = 10)

abunAll <- rbind(abun1, abun2, abun3)

famColSimple_df <- tibble(cols=c(brewer.pal(5,'Set1'), "black"), 
                  family = c('Andrenidae','Apidae','Colletidae',
                           'Halictidae','Megachilidae', "Melittidae"))

famColSimple <- as.character(famColSimple_df$cols)
names(famColSimple) <- famColSimple_df$family

ggplot(abunAll, aes(x = Species, y = Observations, fill = family ))+
  geom_col()+
  scale_fill_manual(values = famColSimple)+
  coord_flip()+
  theme(axis.text.y = element_text(face = "italic"))+
  labs(fill = "Bee Family")+
  facet_wrap(~seas, scales = "free")
```

<!-- ```{r, tab.cap= tcap_topb_1} -->
<!-- kable(slice_head(ecot1 %>% filter(!str_detect(genSpp, "spp\\.")) %>% group_by(genSpp) %>% count() %>%  -->
<!--                    ungroup() %>% arrange(desc(n)) %>% rename(Species = genSpp, Observations = n),  -->
<!--                  n = 10) ) # REPLACED WITH ABOVE FIGURES -->
<!-- ``` -->

<!-- ```{r, tab.cap= tcap_topb_2} -->
<!-- kable(slice_head(ecot2 %>% filter(!str_detect(genSpp, "spp\\.")) %>% group_by(genSpp) %>% count() %>%  -->
<!--                    ungroup() %>% arrange(desc(n)) %>% rename(Species = genSpp, Observations = n),  -->
<!--                  n = 10) ) -->
<!-- ``` -->

<!-- ```{r, tab.cap= tcap_topb_3} -->
<!-- kable(slice_head(ecot3 %>% filter(!str_detect(genSpp, "spp\\.")) %>% group_by(genSpp) %>% count() %>%  -->
<!--                    ungroup() %>% arrange(desc(n)) %>% rename(Species = genSpp, Observations = n),  -->
<!--                  n = 10) ) -->
<!-- ``` -->


```{r, include=FALSE}
# side by side per county? with county highlighted
ecor2highlight <- stateChao_ecoreg %>% filter(name %in% x) %>% 
  mutate(gap = as.character(round(Estimator - Observed), 0))

# make figure caption
fCap <- str_c("Number of detected bee species highlighted for the ", x, " Ecoregion.")
```
\newpage

```{r, include = FALSE} 
ecoQuad <- st_centroid(stateChao_ecoreg) %>% st_join(quadrants, st_within) %>% print(n = 40) %>% 
  select(name, NA_L3NAME, quad) %>% st_drop_geometry()

stateChao_ecoreg <- stateChao_ecoreg %>% left_join(ecoQuad)
whichEcoQuad <- ecor2highlight %>% left_join(ecoQuad) %>% pull(quad)
ecorSubset <- stateChao_ecoreg %>% ungroup() %>% st_filter(., ecor2highlight, .predicate = st_intersects)
```

```{r, fig.height=8, fig.cap= fCap}
ecoSppPlot <- actualStateRich(stateChao_ecoreg)+
  scale_fill_gradient2(low = "white",
                       mid = '#f1eef6',
                       high = '#045a8d')+
  geom_sf(data = ecor2highlight, colour = "green", alpha = 0, size = 1)+
    labs(x = "", y = "")+
  geom_sf(data = datEcoR %>% filter(name %in% x), size = 0.75)+
  geom_sf_text(data = ecor2highlight, aes(label = Observed), size = 4, color = "yellow")

ecoSppPlot

# cowplot::plot_grid(
# actualStateRich(stateChao_ecoreg)+
#   scale_fill_gradient2(low = "white", 
#                        mid = '#f1eef6',
#                        high = '#045a8d')+
#   geom_sf(data = ecor2highlight, colour = "green", alpha = 0, size = 1)+
#     geom_sf_text(data = ecor2highlight, aes(label = Observed), size = 2)+
#     labs(x = "", y = "") , 
# gapStateRich(stateChao_ecoreg)+
#   scale_fill_gradient2(low = "white", 
#                        mid = '#ffffcc',
#                        high = '#bd0026')+
#   geom_sf(data = ecor2highlight, colour = "green", alpha = 0, size = 1)+
#   geom_sf_text(data = ecor2highlight, aes(label = gap), size = 2)+
#   labs(x = "", y = ""),
# 
# nrow = 2
# )
```

\subsection {Uncommon Bee Species for the `r x` Ecoregion}
```{r, include=FALSE}
tcap_unc <- paste0("Bees with less than 15 project-wide observations, and less than 10 in the ",  x, " Ecoregion.")
```


```{r, tab.cap= tcap_unc}
uncft <- flextable(singleEcoNoMissing %>% 
  group_by(genSpp) %>% 
  count() %>% ungroup() %>% 
  semi_join(below15_proj, by = "genSpp") %>% 
    rename(`Bee Species` = genSpp, Observations = n)
  ) %>% 
  italic(j = 1) %>% 
  fontsize(size = 10) %>% 
  align( j = 2, align = "center", part = "all") %>% 
  autofit()

# uncftwids <- dim(uncft)$widths/sum(dim(uncft)$widths)*6.5
# 
# uncft <- width(uncft, width = uncftwids)
uncft
```

#### Known Specialist Bee Species present in the `r x` Ecoregion



Additional specialist bees are likely present, but this information is not always well known. If females of a bee species tend to only collect pollen from a certain species, genus, or family of plant but are not present on this list, the bee may be a yet to be described specialist.  

```{r, tab.cap= str_c("Specialist bees present in ", x, " Ecoregion, with the number of times each species has been observed, its relative rarity, and known floral hosts.")}
library(flextable)

specft <- flextable(singleEcoNoMissing %>% group_by(genSpp) %>% count() %>% ungroup() %>%  
         semi_join(specBeesLookup) %>% left_join(specBeesLookup %>% select(genSpp, Status, `Host plant Family Tribe Genera`)) %>% 
  rename(`Bee Species` = genSpp, Obs. = n, Rarity = Status, `Floral Hosts (Literature - Fowler, 2026)` = `Host plant Family Tribe Genera` )) 

specft  <- autofit(specft)
auto_wid <- dim(specft)$widths/sum(dim(specft)$widths)
auto_wid[1] <- 1.5
auto_wid[2] <- 0.5
auto_wid[3] <- 1
auto_wid[4] <- 3.5 # specify widths - auto isn't working for this case

specft <- width(specft, width = auto_wid)

library(officer)
inBorder <- fp_border(color = "lightgrey", width = 0.5)

specft %>% 
  italic( j = 1) %>% 
  align(j = 2, align = "center", part = "all") %>% 
  fontsize(size = 9, part = "body") %>% 
  fontsize(size = 10, part = "header") %>% 
  border_inner_h(border = inBorder, part = "body") 
```

