---
title: "`r params$collectorName`"
author: "`r params$state` Bee Atlas Collection and Identification Report - `r params$year`"
output: 
  bookdown::pdf_document2:
    toc: false
    fig_caption: yes
    keep_tex: false
params: 
  collectorName: ""
  state: ""
  year: ""
geometry: margin=2cm
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \setlength{\tabcolsep}{0.35em} %Used to adjust cell padding in tables

---

```{r setup, include=FALSE}
#OREGON BEE ATLAS PROJECT SUMMARY SHEETS
#WRITTEN BY SAM ROBINSON, FALL 2020, updated 2025 - Rowan Rampton
library(knitr)
library(kableExtra)
library(tidyverse)
library(ggpubr)
library(RColorBrewer)
library(sf)
library(ggspatial) #library(ggsn) replaced
library(shadowtext)

#knitr options
opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE)

opts_knit$set(eval.after = "fig.cap")
options(dplyr.summarise.inform=FALSE, warn=-1) #Suppress warnings
theme_set(theme_classic())
```


```{r, include=FALSE}
#Load data - needs to be done in shiny app head to work properly
# load("Data/cleanedNew.Rdata")

#Volunteer data only

# volDat <- datS

collStringSTR <- paste0("^", params$collectorName, "$")

collString <- params$collectorName 

singlesR <- singles %>% filter(str_detect(collector, collStringSTR)) # might need to use collStringSTR here

# for now - removing missing obs to see if it'll plot better

stateSTR <- paste0("^", params$state, "$")

stateString <- params$state
```


```{r, include=FALSE}
# DONE - datS = filtered to state, data = spp level IDs, #dataNoNA = spp level in year of report
stateWithin <- usStates %>% select(STUSPS,geometry) %>% rbind(bcBorder %>% select(ABRVN) %>% rename(STUSPS = ABRVN)) %>% 
  st_transform(crs = st_crs(dat)) %>% 
  filter(STUSPS == stateString)
  
datS <- dat %>% filter(str_detect(state, stateString)) %>% 
  st_filter(., stateWithin, .pred = st_within) # getting some points coming in external to OR
  
```


```{r, include=FALSE}
ecoReg <- allEcoReg %>% filter(str_detect(STUSPS, stateSTR))
allCounty <- allCounty %>% filter(str_detect(STUSPS, stateSTR))

stateNameNice <-unique(allCounty$stateName)

regVar <- if_else(stateString %in% "BC", "region", "county")

stProvVar <- if_else(stateString %in% "BC", "province", "state")

regVarPlural <- if_else(stateString %in% "BC", "regions", "counties")

# ecoRegProvVar <- if_else(stateString %in% "BC", "ecoprovince", "ecoregion")
# 
# ecoRegProvVar_plur <- if_else(stateString %in% "BC", "ecoprovinces", "ecoregions")

ecoRegLvlVar <- if_else(stateString %in% "BC", "ecoprovince", "ecoregion (Level III)")

```


```{r, include=FALSE}
data <- datS %>% filter(!is.na(genus)) %>% filter(!is.na(family)) %>% # do pre filtering here - only bees ided to genus or family
  filter(samplingProtocol == "aerial net") %>% # only netted bees
  filter(!is.na(day)) # remove missing dates - only 1 present in 2023 filter(year %in% 2023)

dataNoNA <- data %>% filter(year %in% params$year) # data for 2023 only

#Get top 3 (species richness) collectors
top3Spp <- dataNoNA %>% st_drop_geometry() %>% group_by(collector) %>%
  summarize(n=n(),nSpp=length(unique(genSpp))) %>% 
  arrange(desc(nSpp)) %>% slice(1:3) 

#Get top 3 (number of specimens) collectors
top3N <- dataNoNA %>% st_drop_geometry() %>% count(collector) %>%
  arrange(desc(n)) %>% slice(1:3) 

#Get overall taxonomic accuracy ratings
accuracy23 <- dataNoNA %>% st_drop_geometry() %>% 
  dplyr::select(collector,genusVolDet,genus,volunteerGenSpp,genSpp) %>% 
  mutate(across(c(volunteerGenSpp,genSpp),~ifelse(grepl('spp\\.',.x),NA,.x))) %>% 
  mutate(genusMatch=genusVolDet==genus) %>% 
  mutate(genSppMatch=volunteerGenSpp==genSpp) %>% 
  group_by(collector) %>% 
  summarize(nSpecimens=n(),
    genusID=length(na.omit(genusMatch)),genusCorrect=sum(na.omit(genusMatch)),
    genusAcc=genusCorrect/genusID,
    genSppID=length(na.omit(genSppMatch)),genSppCorrect=sum(na.omit(genSppMatch)),
    genSppAcc=genSppCorrect/genSppID) %>% ungroup() %>% 
  mutate(across(contains('Acc'),~ifelse(is.nan(.x),NA,.x)))

accuracyAll <- data %>% st_drop_geometry() %>% 
  dplyr::select(collector,genusVolDet,genus,volunteerGenSpp,genSpp) %>% 
  mutate(across(c(volunteerGenSpp,genSpp),~ifelse(grepl('spp\\.',.x),NA,.x))) %>% 
  mutate(genusMatch=genusVolDet==genus) %>% 
  mutate(genSppMatch=volunteerGenSpp==genSpp) %>% 
  group_by(collector) %>% 
  summarize(nSpecimens=n(),
    genusID=length(na.omit(genusMatch)),genusCorrect=sum(na.omit(genusMatch)),
    genusAcc=genusCorrect/genusID,
    genSppID=length(na.omit(genSppMatch)),genSppCorrect=sum(na.omit(genSppMatch)),
    genSppAcc=genSppCorrect/genSppID) %>% ungroup() %>% 
  mutate(across(contains('Acc'),~ifelse(is.nan(.x),NA,.x)))

# Linc's ideas for "gold stars"
# # Gold stars for various categories. 
# *most specimens - shown in overall output
# *most species - shown in overall output
# *only species X - this is addressed by the "singles"
# *only genus X - this is addressed by the "singles"
# *most genus X - trickier to do; would have to exclude "singles"
# *most genera - could put into overall output

```

```{r, include= FALSE}
volDat <- dataNoNA %>% filter(str_detect(collector, collStringSTR))

volAccuracy23 <- accuracy23 %>% filter(str_detect(collector, collStringSTR))

volAccuracyAll <- accuracyAll %>% filter(str_detect(collector, collStringSTR))

volDatAllYrs <- data %>% filter(str_detect(collector, collStringSTR)) # for 2nd page of aggregate bees

volIdTBA <- datS %>% filter(str_detect(collector, collStringSTR))

volIdTBA_curYear <- datS %>% filter(year %in% params$year) %>% filter(str_detect(collector, collStringSTR))
```

```{r, include=F}

NsppA <- length(unique(na.omit(volDat$genSpp[!grepl('spp\\.',volDat$genSpp)])))
NgenA <- length(unique(na.omit(volDat$genus)))

#These equations appear to make OK text size adjustments
textSizeA <- c(3.21*log10(NsppA)^(-3.378), 1.26*log10(NgenA)^(-2.22), 1)
textSizeA <- ifelse(textSizeA>1,1,textSizeA) #1 is maximum scaling
textSizeA[3] <- 1.1 #Make family labels larger

```
\section{Your `r params$year` Collections}

`r if_else(nrow(volDat)==0, true = str_c(params$collectorName, " caught ", nrow(volIdTBA_curYear),  if_else(nrow(volIdTBA_curYear)==1, " bee", " bees"), " in ", params$year, if_else(nrow(volIdTBA_curYear)==0, ".", str_c("but none have currently been identified to species. Identifying the ", nrow(datS %>% filter(year%in%params$year)), " bees caught in ", params$year, " takes a long time!"))), false = str_c(params$collectorName, " caught ", nrow(volDat), " bees across ", length(unique(volDat$county)), " ", ifelse(length(unique(volDat$county))>1, regVarPlural, regVar), " from ", format(min(na.omit(volDat$date)),format="%B %d, %Y"), " to ", format(max(na.omit(volDat$date)), format="%B %d, %Y"), " representing ", length(unique(na.omit(volDat$genSpp))), " unique taxa, including ", length(unique(na.omit(volDat$genSpp[!grepl('spp\\.',volDat$genSpp)]))), " unique species."))`


```{r, out.width="100%", fig.cap = paste0("Bees caught by ", collString, ", broken down by species, genus, and family.")}
abundPlots(volDat,colourSet="Set1", scaleYtext = textSizeA)
```

\newpage

\section{All Your Collections}

`r if_else(nrow(volDatAllYrs)==0, true = str_c(params$collectorName, " caught ", nrow(volIdTBA), " bees overall, but none have currently been identified to species. Identifying ", nrow(datS), " bees (and counting) takes a long time!"), false = str_c(params$collectorName, " caught ", nrow(volDatAllYrs), " bees across ",  length(unique(volDatAllYrs$county)), " ", ifelse(length(unique(volDatAllYrs$county))>1, regVarPlural, regVar), " from ",  format(min(na.omit(volDatAllYrs$date)),format="%B %d, %Y"), " to ", format(max(na.omit(volDatAllYrs$date)),format="%B %d, %Y"), ", representing ", length(unique(na.omit(volDatAllYrs$genSpp))), " unique taxa, including ", length(unique(na.omit(volDatAllYrs$genSpp[!grepl('spp\\.',volDatAllYrs$genSpp)]))), " unique species. ", ifelse(nrow(singlesR)>0,paste0(params$collectorName, " also caught the only _",singlesR$genSpp[1],"_ in the collection!"),"")))`


```{r, include=F}

NsppA <- length(unique(na.omit(volDatAllYrs$genSpp[!grepl('spp\\.',volDatAllYrs$genSpp)])))
NgenA <- length(unique(na.omit(volDatAllYrs$genus)))

#These equations appear to make OK text size adjustments
textSizeA <- c(3.21*log10(NsppA)^(-3.378), 1.26*log10(NgenA)^(-2.22), 1)
textSizeA <- ifelse(textSizeA>1,1,textSizeA) #1 is maximum scaling
textSizeA[3] <- 1.1 #Make family labels larger

```

```{r, out.width="100%", fig.cap = paste0("Bees caught by ", collString, ", broken down by species, genus, and family.")}
abundPlots(volDatAllYrs,colourSet="Set1", scaleYtext = textSizeA)
```


```{r volMap, warning=FALSE, fig.cap= paste0("Bee catch locations for ", collString, " (within ", stateNameNice, "), along with total catches per ", regVar ,".")}
volCounties <- volDatAllYrs %>% count(county) %>% 
  st_join(allCounty,.) %>% mutate(n = replace_na(n, 0))

ggplot(volCounties)+
  geom_sf(aes(fill=n),show.legend = FALSE)+
  geom_sf(data=volDatAllYrs,col="black",size=0.5)+
  geom_shadowtext(data=volCounties,
                  aes(x=st_coordinates(st_centroid(volCounties))[,1],y=st_coordinates(st_centroid(volCounties))[,2],
                      label=n))+
  # scale_fill_continuous(low="lightblue",high="red",na.value="white")+
  scale_fill_distiller(palette="OrRd",direction=1,na.value="white")+
  annotation_scale()+ # replaced obsolete ggsn N bar and scale below with ggspatial equivalent
  annotation_north_arrow(location = "tr", style = north_arrow_minimal())+
  #scalebar(allCounty,dist=25,dist_unit="mi",transform=FALSE,model="NAD83",st.size=3,location="bottomright") +
  #north(allCounty,symbol=3, location="topleft")+
  theme_bw() + theme(axis.text=element_blank(), axis.title=element_blank(),
                    axis.line = element_blank(),axis.ticks = element_blank(),
                    panel.grid = element_blank())

```

\newpage

\section{Total Catches}

Volunteers from the `r stateNameNice` Bee Atlas project caught `r nrow(dataNoNA)` bees across `r length(unique(dataNoNA$county))` `r ifelse(length(unique(dataNoNA$county))>1, regVarPlural, regVar)` from  `r format(min(na.omit(dataNoNA$date)),format="%B %d, %Y")` to `r format(max(dataNoNA$date),format="%B %d, %Y")`, representing `r dataNoNA %>% filter(species!="spp.") %>% pull(genSpp) %>% unique() %>% length()` unique species and `r dataNoNA %>% pull(genus) %>% unique() %>% length()` unique genera. The __Nimble Net Kudos__ (most specimens collected) goes to `r top3N$collector[1]`, `r top3N$collector[2]`, and `r top3N$collector[3]`, who caught a total of `r top3N$n[1]`, `r top3N$n[2]`, and `r top3N$n[3]` specimens. The _positive_ kind of __Darwin Award__ (most species collected) goes to `r top3Spp$collector[1]`, `r top3Spp$collector[2]`, and `r top3Spp$collector[3]`, who caught a total of `r top3Spp$nSpp[1]`, `r top3Spp$nSpp[2]`, and `r top3Spp$nSpp[3]` unique species. Well done!

```{r allSppFig, fig.height=11*0.6,fig.width=9*0.6, fig.cap = "Bees caught by all volunteers, broken down by species, genus, and family."}

#Number of unique species and genera
Nspp <- length(unique(na.omit(dataNoNA$genSpp[!grepl('spp\\.',dataNoNA$genSpp)])))
Ngen <- length(unique(na.omit(dataNoNA$genus)))

#These equations appear to make OK text size adjustments
textSize <- c(3.21*log10(Nspp)^(-2.77), 1.26*log10(Ngen)^(-2.12), 1)
textSize <- ifelse(textSize>1,1,textSize) #1 is maximum scaling
textSize[3] <- 1.1 #Make family labels larger

#Originally c(0.35,0.5,1.1)
abundPlots(dataNoNA,colourSet="Set1",scaleYtext = textSize,keepSpp = FALSE)
```

```{r allCatches, fig.height=9,fig.width=8, fig.cap = paste0("Number of bee specimens and unique bee species caught by all volunteers, with your effort shown in red. This graph should give you an idea of how many specimens you would need to catch to begin seeing rarer bee species.")}

#Species accumulation curve 
data %>% st_drop_geometry() %>% 
  filter(!grepl('spp.',genSpp)) %>% 
  group_by(collector) %>% 
  summarize(n=n(),nSpp=length(unique(genSpp))) %>%
  mutate(isCollector=collector==params$collectorName) %>% 
  arrange(isCollector) %>% 
  ggplot(aes(n,nSpp))+
  geom_smooth(method="lm",formula=y~sqrt(x),col='black',fill='black',alpha=0.2)+
  geom_point(aes(col=isCollector),show.legend = FALSE)+
  ylim(0,NA)+
  labs(x="Number of specimens",y="Number of species")+
  scale_colour_manual(values=c('black','red'))

# scaleYtext <- 0.7
# 
# # Bar plot of catches per volunteer, (x = Collector, y = # specimens), with hline and 500 and 50 bees
# p1 <- datS %>% st_drop_geometry() %>% filter(!grepl('NA',collector)) %>% count(collector) %>%
#   arrange(n) %>% mutate(collector=factor(collector,levels=collector)) %>% 
#   mutate(isCol=collector=="VOLUNTEER_NAME") %>% 
#   ggplot(aes(x=collector,y=n))+geom_col(aes(fill=isCol),show.legend=FALSE)+
#   geom_hline(yintercept=c(50,500,1000),linetype="dashed",col="grey")+
#   coord_flip()+scale_fill_manual(values=c("black","red"))+
#   labs(y="Number of Bees",x=NULL)+
#   scale_y_continuous(breaks=c(50,250,500,750,1000,1250))+
#   theme(#axis.text.y=element_text(vjust=0.5,size=8*scaleYtext)
#     axis.text.y=element_blank(),
#     axis.ticks.y = element_blank())
# 
# # Bar plot of species per volunteer, (x = Collector, y = # specimens), with hline and 500 and 50 bees
# p2 <- datS %>% st_drop_geometry() %>% group_by(collector) %>% filter(!grepl('NA',collector)) %>%
#   summarize(n=length(unique(genSpp))) %>% 
#   arrange(n) %>% mutate(collector=factor(collector,levels=collector)) %>% 
#   mutate(isCol=collector=="VOLUNTEER_NAME") %>% 
#   ggplot(aes(x=collector,y=n))+geom_col(aes(fill=isCol),show.legend=FALSE)+
#   geom_hline(yintercept=c(50,100),linetype="dashed",col="grey")+
#   coord_flip()+scale_fill_manual(values=c("black","red"))+
#   labs(y="Number of Species",x=NULL)+
#   scale_y_continuous()+
#   theme(#axis.text.y=element_text(vjust=0.5,size=8*scaleYtext)
#     axis.text.y=element_blank(),
#     axis.ticks.y = element_blank())
# 
# ggarrange(p1,p2,ncol=2)

```

```{r countyBeeMap, fig.cap= str_c("Total specimens caught per ", regVar, ", along with catch location of each specimen (black dots). For genus- and species-specific information for each ", regVar, ", see Tables 3 and 4.")}

#Map of county records
datS %>% ggplot()+ # datS includes specimens not IDed to species, data would be only spp level IDs
  geom_sf(data=allCounty,aes(fill=ifelse(nRecords==0,NA,nRecords)),show.legend = FALSE)+
  geom_sf(size=0.5)+
 # geom_sf_text(data=st_centroid(allCounty),aes(geometry=geometry,label=nRecords),col="white",size=3)+ #Regular text is hard to read
  geom_shadowtext(data=allCounty,
                  aes(x=st_coordinates(st_centroid(allCounty))[,1],y=st_coordinates(st_centroid(allCounty))[,2],
                      label=ifelse(nRecords==0,NA,nRecords)))+
  annotation_scale()+ # replaced obsolete ggsn N bar and scale below with ggspatial equivalent
  annotation_north_arrow(location = "tr", style = north_arrow_minimal())+
  #scalebar(allCounty,dist=25,dist_unit="mi",transform=FALSE,model="NAD83",st.size=3,location="bottomright") +
  #north(allCounty,symbol=4,location="topleft")+
  # scale_fill_continuous(low="lightblue",high="red",na.value="white")+
  scale_fill_distiller(palette="OrRd",direction=1,na.value="white")+
  theme_bw() + theme(axis.text=element_blank(), axis.title=element_blank(),
                     axis.line = element_blank(),axis.ticks = element_blank(),
                     panel.grid = element_blank())

```

```{r ecoRegBeeMap, fig.cap=str_c("Total catches per ", ecoRegLvlVar, ", along with catch location of each specimen (black dots).")}

#Map of ecoregion records
ecoReg %>% 
  mutate(name = str_replace(name, "Klamath Mountains/California High North Coast Range", "Klamath Mountains/\nCalifornia High North \nCoast Range")) %>% # for better figure appearance
  ggplot()+
  geom_sf(aes(fill=name))+
  geom_sf(data=datS,size=0.5)+ # include specimens not IDed to species, use #data if only for spp level IDs
  geom_shadowtext(aes(x=st_coordinates(st_centroid(ecoReg))[,1],y=st_coordinates(st_centroid(ecoReg))[,2],
                      label=nRecords))+
  annotation_scale()+ # replaced obsolete ggsn N bar and scale below with ggspatial equivalent
  annotation_north_arrow(location = "tr", style = north_arrow_minimal())+
  #scalebar(ecoReg,dist=25,dist_unit="mi",transform=FALSE,model="NAD83",st.size=3,location="bottomright") +
  #north(ecoReg,symbol=4,location="topleft")+labs(fill="Ecoregion")+
  scale_fill_brewer(type="qual",palette=3)+
  theme_bw() + theme(axis.text=element_blank(), axis.title=element_blank(),
                     axis.line = element_blank(),axis.ticks = element_blank(),
                     panel.grid = element_blank(), legend.text = element_text(size = 8))

```


\newpage

\section{Flight Phenology}
```{r, eval= params$state %in% c("WA", "OR"), child= "V3_child_phen_cascades.Rmd"}
```

```{r, eval= params$state %in% c("ID", "NM", "CO", "AZ", "BC"), child= "V3_child_phen_non.Rmd"}
```

\newpage

\section{Plant genera}

```{r}
#Get top 3 (flower genera richness) collectors
top3flwCol <- data %>% st_drop_geometry() %>%  group_by(collector) %>%
  summarize(n=n(),nGen=length(unique(genusPlant))) %>% 
  arrange(desc(nGen))

#Get the most popular (most bee specimens) flowers
top3flw <- data %>% st_drop_geometry() %>% filter(!is.na(genusPlant)) %>% count(genusPlant) %>% 
  arrange(desc(n)) %>% slice(1:3)

top3flwSpp <- data %>% st_drop_geometry() %>% filter(!is.na(genusPlant)) %>% group_by(genusPlant) %>% summarize(nSpp=length(unique(genSpp))) %>% 
  arrange(desc(nSpp)) %>% slice(1:3)

```

Volunteers collected specimens from a total of `r length(unique(na.omit(data$genusPlant)))` unique flower genera, with most volunteers sampling from `r median(top3flwCol$nGen)` flower genera (median value). The __Flower Power Kudos__ (most sampled flower genera) goes to `r top3flwCol$collector[1]`, `r top3flwCol$collector[2]`, and `r top3flwCol$collector[3]`, who collected bees from a total of `r top3flwCol$nGen[1]`, `r top3flwCol$nGen[2]`, and `r top3flwCol$nGen[3]` genera of flowers. Well done!

The flower genera that had the most specimens caught on them were _`r top3flw$genusPlant[1]`_, _`r top3flw$genusPlant[2]`_, and _`r top3flw$genusPlant[3]`_, which yielded a total of `r top3flw$n[1]`, `r top3flw$n[2]`, and `r top3flw$n[3]` specimens. The flower genera that were popular with the most species of bees were _`r top3flwSpp$genusPlant[1]`_, _`r top3flwSpp$genusPlant[2]`_, and _`r top3flwSpp$genusPlant[3]`_, hosting a total of `r top3flwSpp$nSpp[1]`, `r top3flwSpp$nSpp[2]`, and `r top3flwSpp$nSpp[3]` unique bee species. See Tables \@ref(tab:flwGenCount) and \@ref(tab:flwGenDiv) for more details.

```{r countyPlantMap, fig.cap=str_c("Recorded number of flower genera per ", regVar, ".")}

#Map of county records
datS %>% ggplot()+
  geom_sf(data=allCounty,aes(fill=ifelse(nFlwGenera==0,NA,nFlwGenera)),show.legend = FALSE)+
  #geom_sf(size=0.5)+ # don't need to show individual bee dots on a genus map
  # geom_sf_text(data=st_centroid(allCounty),aes(geometry=geometry,label=nFlwGenera),col="white",size=3)+
  geom_shadowtext(data=allCounty,
                  aes(x=st_coordinates(st_centroid(allCounty))[,1],y=st_coordinates(st_centroid(allCounty))[,2],
                      label=ifelse(nFlwGenera==0,NA,nFlwGenera)))+
  annotation_scale()+ # replaced obsolete ggsn N bar and scale below with ggspatial equivalent
  annotation_north_arrow(location = "tr", style = north_arrow_minimal())+
  #scalebar(allCounty,dist=25,dist_unit="mi",transform=FALSE,model="NAD83",st.size=3,location="bottomright") +
  #north(allCounty,symbol=4,location="topleft")+
  # scale_fill_continuous(low="lightblue",high="red",na.value="white")+
  scale_fill_distiller(palette="Greens",direction=1,na.value="white")+
  theme_bw() + theme(axis.text=element_blank(), axis.title=element_blank(),
                     axis.line = element_blank(),axis.ticks = element_blank(),
                     panel.grid = element_blank())

```

```{r flwGenCount}

data %>% st_drop_geometry() %>% 
  filter(!is.na(genusPlant)) %>% count(genusPlant,name='N') %>% arrange(N) %>% 
  wideTable(countBy=genusPlant,countNum=N,nsplit=7,colNames=c("Genus","Count"),
          caption='Number of bee specimens collected from each plant genus. Plants with few records are great targets for future sampling.',font_size=4.15)

```


```{r flwGenDiv}
# Ideally this would be rarefied richness (adjust for abundance), but OK for now
 data %>% st_drop_geometry() %>% filter(!is.na(genusPlant)) %>% 
  group_by(genusPlant) %>% summarize(nSpp=length(unique(genSpp))) %>% arrange(nSpp) %>%
  wideTable(countBy=genusPlant,countNum=nSpp,nsplit=7,colNames=c("Genus","Count"),
          caption='Number of bee species collected from each plant genus',font_size=4.15)
  
```

\begin{landscape}

\section{`r paste0(str_to_title(regVar)," records")` }

```{r genusCountyTab}

data %>% st_drop_geometry() %>%
  abundTable(county,genus,TRUE) %>%
  kable(format="latex",longtable=TRUE,caption=str_c("Number of bee specimens from each ", regVar, ", by genus. You may want to focus your sampling in under-sampled counties.")) %>%
  kable_styling(latex_options = c("striped","repeat_header"),font_size=4.25) %>%
  row_spec(0, angle=90) %>% #Rotate row 0 (header) by 90 degrees
  column_spec(column=1,italic=TRUE,latex_column_spec = "r") %>% #Genus names
  column_spec(column=2:length(unique(data$county))+1) %>% #Counties + data
  column_spec(column=length(unique(data$county))+2,latex_column_spec = "l")

```

\newpage

```{r sppCountyTab}

nCounties <- data %>% 
    group_by(county, species) %>% 
    count() %>% 
    filter(!is.na(species)) # make column specs robust to counties without at least one bee ided to species

data %>% st_drop_geometry() %>% filter(species!="spp.") %>%
  abundTable(county,genSpp,TRUE) %>%
  kable(format="latex",longtable=TRUE,align=paste(c(rep('r',nrow(.)-1),'l'),collapse=''),
        caption=str_c("Number of bee specimens from each ", regVar ,", by species")) %>%
  kable_styling(latex_options = c("striped","repeat_header"),font_size=4) %>%
  row_spec(0, angle=90) %>% #Rotate row 0 (header) by 90 degrees
  ## TODO using column spec here for some reason doesn't work, looks okay ish for now but look closer in later versions
  column_spec(column=1,italic=TRUE,latex_column_spec = 'r') %>% #Spp names
  column_spec(column=2:length(unique(nCounties$county))+1) %>% #Counties + data
  column_spec(column=length(unique(nCounties$county))+2,latex_column_spec = 'l')  #TOTAL column at the end

```

\end{landscape}

\newpage

\section{Taxonomic Accuracy, `r params$year`}

In `r params$year`, `r ifelse(nrow(volDat)<1, yes = str_c(" you did not have specimens that were identified by the ", stateNameNice, " Bee Atlas"), no = str_c(" you identified ", volAccuracy23$genusID, " of your ",volAccuracy23$nSpecimens, " specimens to genus level and ", volAccuracy23$genSppID, " to species level ", with(volAccuracy23,ifelse(genusID>0,paste0(", with a genus-level identification accuracy of ",round(100*genusAcc,1),"%"),"")), with(volAccuracy23,ifelse(genSppID>0,paste0(" and a species-level identification accuracy of ",round(100*genSppAcc,1),"%"),"")) ) )` (see Table \@ref(tab:volDetAccuracy23)).
In total, volunteers from the `r stateNameNice` Bee Atlas project identified `r round(100*with(accuracy23,sum(genusID)/sum(nSpecimens)),1)` % (`r sum(accuracy23$genusID)`) of the `r sum(accuracy23$nSpecimens)` bee specimens to the level of genus, with an average accuracy of `r round(100*with(accuracy23,sum(genusCorrect)/sum(genusID)),1)`%. 
`r if_else(is.nan(round(100*with(accuracy23,sum(genSppCorrect)/sum(genSppID)),1)), "Volunteer identifications to species level did not occur or are not included in the data ", str_c("Volunteers also identified ", round(100*with(accuracy23,sum(genSppID)/sum(nSpecimens)),1), "% (", sum(accuracy23$genSppID), ") of the specimens to species level, and had an average accuracy of ",round(100*with(accuracy23,sum(genSppCorrect)/sum(genSppID)),1), "% " )  )` (see Table \@ref(tab:detAccuracy23))`r if_else(is.nan(round(100*with(accuracy23,sum(genSppCorrect)/sum(genSppID)),1)), ".", ". Nicely done!") `

```{r volDetAccuracy23}
#Specific volunteer determination accuracy
accFun <- function(d,vFam,fam,addTotal=TRUE){ #Calculate determination accuracy, using volunteer + actual columns
  require(dplyr)
  d <- d %>% st_drop_geometry() %>% 
    dplyr::select({{vFam}},{{fam}}) %>% 
    filter(!is.na({{vFam}})) %>%
    filter(!is.na({{fam}})) %>% 
    mutate(detCorrect={{vFam}}=={{fam}}) %>% 
    group_by({{fam}}) %>% 
    summarize(n=n(),nCorrect=sum(detCorrect),propCorrect=100*round(nCorrect/n,3)) %>% 
    ungroup()
  if(addTotal){ #Adds "TOTAL" metrics at the end if addTotal==TRUE
    totalN <- sum(pull(d,n)) #Get statistics
    totalNcorrect <- sum(pull(d,nCorrect))
    totalProp <- 100*round(totalNcorrect/totalN,3)
    total <- data.frame('TOTAL',totalN,totalNcorrect,totalProp) #Single-row dataframe
    colnames(total)=colnames(d) #Change column labels
    d <- bind_rows(d,total)
  } 
  return(d)
}

t1 <- volDat %>% accFun(familyVolDet,family) #Family-level accuracy
t2 <- volDat %>% accFun(genusVolDet,genus) #Genus-level
t3 <- volDat %>% filter(!is.na(species),!is.na(specificEpithetVolDet)) %>% 
  accFun(genSpp,volunteerGenSpp) #Species-level
colnames(t3) = colnames(t2) = colnames(t1) #Column names

overallDet <- bind_rows(t1,t2,t3) %>% filter(!is.nan(propCorrect)) 

if(nrow(overallDet)>0){ #If 
  tabBreaks <- c('Family'=nrow(t1),'Genus'=nrow(t2),'Species'=nrow(t3))
  tabBreaks <- tabBreaks[tabBreaks>1] #Gets rid of non-existant classes
  
  kable(overallDet,format="latex",longtable=TRUE,align=c('r','r','r','l'),
        caption=paste0("Your determination accuracy in ", params$year, "."),
        col.names=c('Taxon','Specimens ID-ed','Correct ID','% Correct')) %>% 
    pack_rows(index=tabBreaks, #Break up table by taxa
              background = 'lightgray') %>% 
    column_spec(1,italic=TRUE) %>% 
    kable_styling(latex_options = c("repeat_header"))
} else {
  data.frame('Taxon'='No specimens identified') %>% 
    kable(overallDet,format="latex",longtable=FALSE,align=c('c'),
        caption=paste0("Your determination accuracy in ", params$year, "."))
}

 
```


```{r detAccuracy23}
#Overall volunteer determination accuracy

t1 <- accFun(dataNoNA,familyVolDet,family) #Family-level accuracy
t2 <- accFun(dataNoNA,genusVolDet,genus) #Genus-level
t3 <- dataNoNA %>% filter(!is.na(species),!is.na(specificEpithetVolDet)) %>% #Species-level
  accFun(genSpp,volunteerGenSpp)
colnames(t3) = colnames(t2) = colnames(t1) #Column names

bind_rows(t1,t2,t3) %>% 
  kable(format="latex",longtable=TRUE,align=c('r','r','r','l'),caption= paste0("Determination accuracy for all volunteers in ", params$year, "."),
        col.names=c('Taxon','Specimens ID-ed','Correct ID','% Correct')) %>% 
  pack_rows(index=c('Family'=nrow(t1),'Genus'=nrow(t2),'Species'=nrow(t3)), #Break up table by taxa
            background = 'lightgray') %>% 
    column_spec(1,italic=TRUE) %>% 
  kable_styling(latex_options = c("repeat_header"))
 
```

\newpage

\section{Taxonomic Accuracy, All Years}

`r ifelse(nrow(volDatAllYrs) < 1, yes = str_c("Your collected specimens have not yet been identified by the ", stateNameNice, " Bee Atlas"), no = str_c("Over your time in the Atlas you identified ", volAccuracyAll$genusID, " of your ", volAccuracyAll$nSpecimens, " specimens to genus level and ", volAccuracyAll$genSppID, " to species level ", with(volAccuracyAll,ifelse(genusID>0,paste0(", with a genus-level identification accuracy of ",round(100*genusAcc,1),"%"),"")), " ", with(volAccuracyAll,ifelse(genSppID>0,paste0(" and a species-level identification accuracy of ",round(100*genSppAcc,1),"%"),""))  ))` (see Table \@ref(tab:volDetAccuracyAll)).
In total, volunteers from the `r stateNameNice` Bee Atlas project identified `r round(100*with(accuracyAll,sum(genusID)/sum(nSpecimens)),1)` % (`r sum(accuracyAll$genusID)`) of the `r sum(accuracyAll$nSpecimens)` bee specimens to the level of genus, with an average accuracy of `r round(100*with(accuracyAll,sum(genusCorrect)/sum(genusID)),1)`%. `r if_else(is.nan(sum(accuracyAll$genSppCorrect)/sum(accuracyAll$genSppID)), "Volunteer identifications to species level did not occur or are not included in the data", str_c("Volunteers also identified ", round(100*with(accuracyAll,sum(genSppID)/sum(nSpecimens)),1), "% (", sum(accuracyAll$genSppID), ") of the specimens to species level and had an average accuracy of", round(100*with(accuracyAll,sum(genSppCorrect)/sum(genSppID)),1), "%"))` (See Table \@ref(tab:detAccuracyAll))`r if_else(is.nan(with(accuracyAll,sum(genSppCorrect)/sum(genSppID))), ".", ". Nicely done!") `

```{r volDetAccuracyAll}
#Specific volunteer determination accuracy
accFun <- function(d,vFam,fam,addTotal=TRUE){ #Calculate determination accuracy, using volunteer + actual columns
  require(dplyr)
  d <- d %>% st_drop_geometry() %>% 
    dplyr::select({{vFam}},{{fam}}) %>% 
    filter(!is.na({{vFam}})) %>%
    filter(!is.na({{fam}})) %>% 
    mutate(detCorrect={{vFam}}=={{fam}}) %>% 
    group_by({{fam}}) %>% 
    summarize(n=n(),nCorrect=sum(detCorrect),propCorrect=100*round(nCorrect/n,3)) %>% 
    ungroup()
  if(addTotal){ #Adds "TOTAL" metrics at the end if addTotal==TRUE
    totalN <- sum(pull(d,n)) #Get statistics
    totalNcorrect <- sum(pull(d,nCorrect))
    totalProp <- 100*round(totalNcorrect/totalN,3)
    total <- data.frame('TOTAL',totalN,totalNcorrect,totalProp) #Single-row dataframe
    colnames(total)=colnames(d) #Change column labels
    d <- bind_rows(d,total)
  } 
  return(d)
}

t1 <- volDatAllYrs %>% accFun(familyVolDet,family) #Family-level accuracy
t2 <- volDatAllYrs %>% accFun(genusVolDet,genus) #Genus-level
t3 <- volDatAllYrs %>% filter(!is.na(species),!is.na(specificEpithetVolDet)) %>% 
  accFun(genSpp,volunteerGenSpp) #Species-level
colnames(t3) = colnames(t2) = colnames(t1) #Column names

overallDet <- bind_rows(t1,t2,t3) %>% filter(!is.nan(propCorrect)) 

if(nrow(overallDet)>0){ #If 
  tabBreaks <- c('Family'=nrow(t1),'Genus'=nrow(t2),'Species'=nrow(t3))
  tabBreaks <- tabBreaks[tabBreaks>1] #Gets rid of non-existant classes
  
  kable(overallDet,format="latex",longtable=TRUE,align=c('r','r','r','l'),
        caption="Your determination accuracy.",
        col.names=c('Taxon','Specimens ID-ed','Correct ID','% Correct')) %>% 
    pack_rows(index=tabBreaks, #Break up table by taxa
              background = 'lightgray') %>% 
    column_spec(1,italic=TRUE) %>% 
    kable_styling(latex_options = c("repeat_header"))
} else {
  data.frame('Taxon'='No specimens identified') %>% 
    kable(overallDet,format="latex",longtable=FALSE,align=c('c'),
        caption="Your determination accuracy.")
}

 
```


```{r detAccuracyAll}
#Overall volunteer determination accuracy

t1 <- accFun(data,familyVolDet,family) #Family-level accuracy
t2 <- accFun(data,genusVolDet,genus) #Genus-level
t3 <- data %>% filter(!is.na(species),!is.na(specificEpithetVolDet)) %>% #Species-level
  accFun(genSpp,volunteerGenSpp)
colnames(t3) = colnames(t2) = colnames(t1) #Column names

bind_rows(t1,t2,t3) %>% 
  kable(format="latex",longtable=TRUE,align=c('r','r','r','l'),caption="Determination accuracy for all volunteers.",
        col.names=c('Taxon','Specimens ID-ed','Correct ID','% Correct')) %>% 
  pack_rows(index=c('Family'=nrow(t1),'Genus'=nrow(t2),'Species'=nrow(t3)), #Break up table by taxa
            background = 'lightgray') %>% 
    column_spec(1,italic=TRUE) %>% 
  kable_styling(latex_options = c("repeat_header"))
 
```

```{r}
options(warn=0) #Turn warnings back on
```
