---
title: "VOLUNTEER_NAME Summary Sheet"
author: "Oregon Bee Atlas"
output: 
  pdf_document:
     fig_caption: yes
     keep_tex: false
geometry: margin=2cm
header-includes: 
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  
---

```{r setup, include=FALSE}
#OREGON BEE ATLAS PROJECT SUMMARY SHEETS
#WRITTEN BY SAM ROBINSON, FALL 2020
library(knitr)
library(kableExtra)
library(tidyverse)
library(ggpubr)
library(RColorBrewer)
library(sf)
library(ggsn)
library(shadowtext)

#knitr options
opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE)

options(dplyr.summarise.inform=FALSE, warn=-1) #Suppress warnings
theme_set(theme_classic())

#Load data
setwd("~/Documents/Andony Melathopolous/Volunteer Summary Sheets")
load("./Data/cleaned.Rdata")

#Volunteer data only
volDat <- dat %>% filter(collector=="VOLUNTEER_NAME") 
  
singles <- filter(singles,collector=="VOLUNTEER_NAME")

#Get top 3 (species richness) collectors
top3Spp <- dat %>% st_drop_geometry() %>%  group_by(collector) %>%
  summarize(n=n(),nSpp=length(unique(genSpp))) %>% 
  arrange(desc(nSpp)) %>% slice(1:3) 

#Get top 3 (number of specimens) collectors
top3N <- dat %>%  st_drop_geometry() %>% count(collector) %>%
  arrange(desc(n)) %>% slice(1:3) 

# Linc's ideas for "gold stars"
# # Gold stars for various categories. 
# *most specimens - shown in overall output
# *most species - shown in overall output
# *only species X - this is addressed by the "singles"
# *only genus X - this is addressed by the "singles"
# *most genus X - trickier to do; would have to exclude "singles"
# *most genera - could put into overall output

```

\section{Volunteer Catches}

VOLUNTEER_NAME caught `r nrow(volDat)` bees across `r length(unique(volDat$county))` `r ifelse(length(unique(volDat$county))>1,"counties","county")` from  `r format(min(volDat$date),format="%B %d, %Y")` to `r format(max(volDat$date),format="%B %d, %Y")`, representing `r length(unique(volDat$genSpp))` unique species. `r ifelse(nrow(singles)>0,paste0("VOLUNTEER_NAME also caught the only _",singles$genSpp[1],"_ in the collection!"),"")`

```{r, out.width="100%", fig.cap = "Bees caught by VOLUNTEER_ESCNAME, broken down by species, genus, and family."}
abundPlots(volDat,colSet="Set1")
```

```{r volMap, warning=FALSE, fig.cap="Bee catch locations for VOLUNTEER_ESCNAME, along with total catches per county."}
volCounties <- volDat %>% count(county) %>% 
  st_join(orCounties,.) 
ggplot(volCounties)+
  geom_sf(aes(fill=n),show.legend = FALSE)+
  geom_sf(data=volDat,col="black",size=0.5)+
  geom_shadowtext(data=volCounties,
                  aes(x=st_coordinates(st_centroid(volCounties))[,1],y=st_coordinates(st_centroid(volCounties))[,2],
                      label=n))+
  # scale_fill_continuous(low="lightblue",high="red",na.value="white")+
  scale_fill_distiller(palette="OrRd",direction=1,na.value="white")+
  scalebar(orCounties,dist=25,dist_unit="mi",transform=FALSE,model="NAD83",st.size=3,location="bottomright") +
  north(orCounties,symbol=3, location="topleft")+
  theme_bw() + theme(axis.text=element_blank(), axis.title=element_blank(),
                    axis.line = element_blank(),axis.ticks = element_blank(),
                    panel.grid = element_blank())

```

\newpage

\section{Total Catches}

Volunteers from the Oregon Bee Atlas project caught `r nrow(dat)` bees across `r length(unique(dat$county))` `r ifelse(length(unique(dat$county))>1,"counties","county")` from  `r format(min(dat$date),format="%B %d, %Y")` to `r format(max(dat$date),format="%B %d, %Y")`, representing `r dat %>% filter(species!="spp.") %>% pull(genSpp) %>% unique() %>% length()` unique species and `r dat %>% filter(species!="spp.") %>% pull(genus) %>% unique() %>% length()` unique genera. The __Nimble Net Award__ (most specimens collected) goes to `r top3N$collector[1]`, `r top3N$collector[2]`, and `r top3N$collector[3]`, who caught a total of `r top3N$n[1]`, `r top3N$n[2]`, and `r top3N$n[3]` specimens. The _positive_ kind of __Darwin Award__ (most species collected) goes to `r top3Spp$collector[1]`, `r top3Spp$collector[2]`, and `r top3Spp$collector[3]`, who caught a total of `r top3Spp$nSpp[1]`, `r top3Spp$nSpp[2]`, and `r top3Spp$nSpp[3]` unique species. Well done!


```{r, fig.height=11*0.6,fig.width=9*0.6, fig.cap = "Bees caught by all volunteers, broken down by species, genus, and family."}
abundPlots(dat,colSet="Set1",scaleYtext = c(0.4,0.5,1.1))
```

```{r, fig.height=9,fig.width=8, fig.cap = "Number of bee specimens and number of unique bee species caught by all volunteers."}

# Bar plot of catches per volunteer, (x = Collector, y = # specimens), with hline and 500 and 50 bees
p1 <- dat %>% st_drop_geometry() %>%  count(collector) %>% 
  arrange(n) %>% mutate(collector=factor(collector,levels=collector)) %>% 
  mutate(isCol=collector=="VOLUNTEER_NAME") %>% 
  ggplot(aes(x=collector,y=n))+geom_col(aes(fill=isCol),show.legend=FALSE)+
  geom_hline(yintercept=c(50,500,1000),linetype="dashed",col="grey")+
  coord_flip()+scale_fill_manual(values=c("black","red"))+
  labs(y="Number of Specimens",x=NULL)+
  scale_y_continuous(breaks=c(50,250,500,750,1000,1250))

# Bar plot of species per volunteer, (x = Collector, y = # specimens), with hline and 500 and 50 bees
p2 <- dat %>% st_drop_geometry() %>%  group_by(collector) %>% summarize(n=length(unique(genSpp))) %>% 
  arrange(n) %>% mutate(collector=factor(collector,levels=collector)) %>% 
  mutate(isCol=collector=="VOLUNTEER_NAME") %>% 
  ggplot(aes(x=collector,y=n))+geom_col(aes(fill=isCol),show.legend=FALSE)+
  geom_hline(yintercept=c(50,100),linetype="dashed",col="grey")+
  coord_flip()+scale_fill_manual(values=c("black","red"))+
  labs(y="Number of Unique Species",x=NULL)+
  scale_y_continuous()

ggarrange(p1,p2,ncol=2)

```

```{r countyBeeMap, fig.cap="Total specimens caught per county, along with catch location of each specimen (black dots). For genus- and species-specific information for each county, see Tables 3 and 4."}

#Map of county records
dat %>% ggplot()+
  geom_sf(data=orCounties,aes(fill=ifelse(nRecords==0,NA,nRecords)),show.legend = FALSE)+
  geom_sf(size=0.5)+
  # geom_sf_text(data=st_centroid(orCounties),aes(geometry=geometry,label=nRecords),col="white",size=3)+ #Regular text is hard to read
  geom_shadowtext(data=orCounties,
                  aes(x=st_coordinates(st_centroid(orCounties))[,1],y=st_coordinates(st_centroid(orCounties))[,2],
                      label=ifelse(nRecords==0,NA,nRecords)))+
  scalebar(orCounties,dist=25,dist_unit="mi",transform=FALSE,model="NAD83",st.size=3,location="bottomright") +
  north(orCounties,symbol=4,location="topleft")+
  # scale_fill_continuous(low="lightblue",high="red",na.value="white")+
  scale_fill_distiller(palette="OrRd",direction=1,na.value="white")+
  theme_bw() + theme(axis.text=element_blank(), axis.title=element_blank(),
                     axis.line = element_blank(),axis.ticks = element_blank(),
                     panel.grid = element_blank())

```

```{r ecoRegBeeMap, fig.cap="Total catches per (Level III) ecoregion, along with catch location of each specimen (black dots)."}

#Map of ecoregion records
orEcoReg %>% ggplot()+
  geom_sf(aes(fill=name))+
  geom_sf(data=dat,size=0.5)+
  geom_shadowtext(aes(x=st_coordinates(st_centroid(orEcoReg))[,1],y=st_coordinates(st_centroid(orEcoReg))[,2],
                      label=nRecords))+
  scalebar(orEcoReg,dist=25,dist_unit="mi",transform=FALSE,model="NAD83",st.size=3,location="bottomright") +
  north(orEcoReg,symbol=4,location="topleft")+labs(fill="Ecoregion")+
  scale_fill_brewer(type="qual",palette=3)+
  theme_bw() + theme(axis.text=element_blank(), axis.title=element_blank(),
                     axis.line = element_blank(),axis.ticks = element_blank(),
                     panel.grid = element_blank())

```


\newpage

\section{Flight Phenology}

```{r}

quantFun <- function(x, q) { #"Borrowed" from dplyr 1.0.0 summary
  tibble(x = quantile(x, q), q = q)
}

#Info for flight season
phenoRange <- dat %>% st_drop_geometry() %>% mutate(doy=as.numeric(format(date,format="%j"))) %>%
  summarize(quantFun(x=doy,q=c(0.05,0.25,0.5,0.75,0.95))) %>% 
  mutate(xOffset=as.Date(as.character(x-10),format="%j"), #x-Location for text label
         yOffset=length(unique(dat$genSpp[!grepl(".spp",dat$genSpp)])), #y-location for text label
         x=as.Date(as.character(x),format="%j"), #Actual date
         lab=paste0(q*100,"%"), # Percentile Label
         dat=format(x,format="%B %d")) #Date label

```

Most bees (90%) were caught between `r phenoRange$dat[1]` and `r phenoRange$dat[5]`, but the peak of season (50% of specimens) was from `r phenoRange$dat[2]` to `r phenoRange$dat[4]`.

```{r sppPhenology, fig.cap="Phenology plot for all bee species, sorted by median abundance times. Percentiles of overall emergence times (50th & 90th) are shown in grey shaded regions. Date ranges for each species (minimum, first, second, third quartiles, and maximum) are shown only for species with >10 specimens.", fig.height=11*0.6, fig.width=10*0.7}

rectDat <- with(phenoRange,data.frame(perc=c("50th","90th"),
                                      ymin=x[q==0.25|q==0.05],ymax=x[q==0.75|q==0.95],
                                      xmin=c(0,0),xmax=rep(yOffset[1],2)))

 
dat %>% st_drop_geometry() %>% mutate(doy=as.numeric(format(date,format="%j"))) %>% 
  filter(species!="spp.") %>% group_by(genSpp) %>% 
  summarize(n=n(),quantFun(x=doy,q=seq(0,1,0.25))) %>%
  mutate(q=rank(q)) %>% 
  pivot_wider(names_from=q,names_prefix="q",values_from=x) %>% 
  mutate(across(c(q1,q2,q4,q5),~ifelse(n>10,.,NA))) %>% 
  ungroup() %>% 
  arrange(desc(q3)) %>% mutate(across(c(q1:q5),~as.Date(as.character(.),format="%j"))) %>% 
  # filter(n>20) %>% 
  mutate(genSpp=factor(genSpp,levels=genSpp)) %>%
  ggplot(aes(x=genSpp)) + geom_linerange(aes(ymin=q1,ymax=q5))+
  
  geom_linerange(aes(ymin=q2,ymax=q4),size=1)+ #Bee emergence dates
  geom_point(aes(y=q3),size=1)+
  
  # geom_text(data=phenoRange,aes(x=yOffset,y=xOffset,label=lab),col="red",alpha=0.8)+ #Percentile markers
  # geom_hline(data=phenoRange,aes(yintercept=x),col="red",linetype="dashed",alpha=0.8)+
  geom_rect(data=rectDat,aes(x=NULL,xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax),alpha=0.2,show.legend = FALSE)+
  
  coord_flip()+labs(x=NULL,y=NULL)+
  theme(panel.grid.major.x=element_line(colour="grey"),
        #panel.grid.minor.x=element_line(colour="grey"),
        axis.text.x=element_text(size=7),
        axis.text.y=element_text(size=3))+
  scale_y_date(date_breaks="1 month",date_labels="%B",date_minor_breaks="1 weeks")
  

```

\newpage

\section{Plant genera}

```{r}
#Get top 3 (flower genera richness) collectors
top3flwCol <- dat %>% st_drop_geometry() %>%  group_by(collector) %>%
  summarize(n=n(),nGen=length(unique(flwGenus))) %>% 
  arrange(desc(nGen))

#Get the most popular (most bee specimens) flowers
top3flw <- dat %>% st_drop_geometry() %>% filter(!is.na(flwGenus)) %>% count(flwGenus) %>% 
  arrange(desc(n)) %>% slice(1:3)

top3flwSpp <- dat %>% st_drop_geometry() %>% filter(!is.na(flwGenus)) %>% group_by(flwGenus) %>% summarize(nSpp=length(unique(genSpp))) %>% 
  arrange(desc(nSpp)) %>% slice(1:3)

```

Volunteers collected specimens from a total of `r length(unique(dat$flwGenus))` unique flower genera, with most volunteers sampling from `r median(top3flwCol$nGen)` flower genera (median value). The __Flower Power Award__ (most sampled flower genera) goes to `r top3flwCol$collector[1]`, `r top3flwCol$collector[2]`, and `r top3flwCol$collector[3]`, who collected bees from a total of `r top3flwCol$nGen[1]`, `r top3flwCol$nGen[2]`, and `r top3flwCol$nGen[3]` genera of flowers. Well done!

The flower genera that had the most specimens caught on them were _`r top3flw$flwGenus[1]`_, _`r top3flw$flwGenus[2]`_, and _`r top3flw$flwGenus[3]`_, which yielded a total of `r top3flw$n[1]`, `r top3flw$n[2]`, and `r top3flw$n[3]` specimens. The flower genera that were popular with the most speces of bees were _`r top3flwSpp$flwGenus[1]`_, _`r top3flwSpp$flwGenus[2]`_, and _`r top3flwSpp$flwGenus[3]`_, hosting at total of `r top3flwSpp$nSpp[1]`, `r top3flwSpp$nSpp[2]`, and `r top3flwSpp$nSpp[3]` unique bee species. See Tables \@ref(tab:flwGenCount) and \@ref(tab:flwGenDiv) for more details.


```{r countyPlantMap, fig.cap="Recorded number of flower genera per county."}

#Map of county records
dat %>% ggplot()+
  geom_sf(data=orCounties,aes(fill=ifelse(nFlwGenera==0,NA,nFlwGenera)),show.legend = FALSE)+
  geom_sf(size=0.5)+
  # geom_sf_text(data=st_centroid(orCounties),aes(geometry=geometry,label=nFlwGenera),col="white",size=3)+
  geom_shadowtext(data=orCounties,
                  aes(x=st_coordinates(st_centroid(orCounties))[,1],y=st_coordinates(st_centroid(orCounties))[,2],
                      label=ifelse(nFlwGenera==0,NA,nFlwGenera)))+
  scalebar(orCounties,dist=25,dist_unit="mi",transform=FALSE,model="NAD83",st.size=3,location="bottomright") +
  north(orCounties,symbol=4,location="topleft")+
  # scale_fill_continuous(low="lightblue",high="red",na.value="white")+
  scale_fill_distiller(palette="Greens",direction=1,na.value="white")+
  theme_bw() + theme(axis.text=element_blank(), axis.title=element_blank(),
                     axis.line = element_blank(),axis.ticks = element_blank(),
                     panel.grid = element_blank())

```

```{r flwGenCount}
ncols <- 4 #Columns to divide into

# #Figure - this works, but is hard to read, and you can't make out any real numbers
# dat %>% st_drop_geometry() %>% filter(!is.na(flwGenus)) %>% count(flwGenus) %>% 
#   arrange(n) %>% mutate(flwGenus=factor(flwGenus,levels=flwGenus)) %>%
#   mutate(ord=cut(rev(rank(as.numeric(flwGenus))),ncols,labels=1:ncols)) %>% 
#   arrange(ord,desc(n)) %>%
#   ggplot(aes(x=flwGenus,y=n))+geom_col() + facet_wrap(~ord,ncol=ncols,scales="free_y")+
#   coord_flip()+labs(y="Number of specimens",x="Flower genus")+
#   theme(axis.text.y=element_text(size=5),axis.text.x=element_text(size=6),
#         strip.background=element_blank(),strip.text=element_blank())

#Table - looks OK, but should probably be made into a function later to avoid replication
dat %>% st_drop_geometry() %>% filter(!is.na(flwGenus)) %>% count(flwGenus) %>% 
  arrange(n) %>% mutate(flwGenus=factor(flwGenus,levels=flwGenus)) %>%
  mutate(ord=cut(rev(rank(as.numeric(flwGenus))),ncols,labels=1:ncols)) %>%
  arrange(ord,desc(n)) %>% mutate(across(c(flwGenus,n),as.character)) %>% 
  group_by(ord) %>% mutate(id=1:n()) %>% ungroup() %>%
  pivot_longer(cols=flwGenus:n) %>% 
  unite(name,name,ord,sep="_") %>% pivot_wider(id_cols=id,names_from=name,values_from=value) %>% 
  select(-id) %>% mutate(across(everything(),~ifelse(is.na(.)," ",.))) %>%
  kable(col.names=rep(c("Genus","Count"),ncols),align=rep(c("r","l"),ncols),
        caption="Number of bee specimens collected from each plant genus") %>% 
  column_spec(column=seq(1,ncols*2,2),italic=TRUE) %>% #Genus names
  kable_styling(latex_options = c("striped"),font_size=6) 
```

```{r flwGenDiv}
ncols <- 4 #Columns to divide into

#Table - ideally this would be rarefied richness (adjust for abundance), but OK for now
dat %>% st_drop_geometry() %>% filter(!is.na(flwGenus)) %>% 
  group_by(flwGenus) %>% summarize(nSpp=length(unique(genSpp))) %>% 
  arrange(nSpp) %>% mutate(flwGenus=factor(flwGenus,levels=flwGenus)) %>%
  mutate(ord=cut(rev(rank(as.numeric(flwGenus))),ncols,labels=1:ncols)) %>%
  arrange(ord,desc(nSpp)) %>% mutate(across(c(flwGenus,nSpp),as.character)) %>% 
  group_by(ord) %>% mutate(id=1:n()) %>% ungroup() %>%
  pivot_longer(cols=flwGenus:nSpp) %>% 
  unite(name,name,ord,sep="_") %>% pivot_wider(id_cols=id,names_from=name,values_from=value) %>% 
  select(-id) %>% mutate(across(everything(),~ifelse(is.na(.)," ",.))) %>%
  kable(col.names=rep(c("Genus","Count"),ncols),align=rep(c("r","l"),ncols),
        caption="Number of unique bee species collected from each plant genus") %>% 
  column_spec(column=seq(1,ncols*2,2),italic=TRUE) %>% #Genus names
  kable_styling(latex_options = c("striped"),font_size=6) 
```


\begin{landscape}

\section{County records}

```{r genusCountyTab}

dat %>% st_drop_geometry() %>% 
  abundTable(county,genus,TRUE) %>%
  kable(format="latex",longtable=TRUE,caption="Number of bee specimens from each county, by genus") %>% row_spec(0, angle=90) %>%
  column_spec(column=1,italic=TRUE) %>% #Genus names
  column_spec(column=2:length(unique(dat$county))+1,width="0.6em") %>% #Counties
  column_spec(column=length(unique(dat$county))+2,width="2em") %>% #Total
  kable_styling(latex_options = c("striped","repeat_header"),font_size=6) 
  
```

\newpage

```{r sppCountyTab}

dat %>% st_drop_geometry() %>% filter(species!="spp.") %>%  
  abundTable(county,genSpp,TRUE) %>%
  kable(format="latex",longtable=TRUE,caption="Number of bee specimens from each county, by species") %>% row_spec(0, angle=90) %>%
  column_spec(column=1,italic=TRUE) %>% #Spp names
  column_spec(column=2:length(unique(dat$county))+1,width="0.6em") %>% #Counties
  column_spec(column=length(unique(dat$county))+2,width="2em") %>% #Total
  kable_styling(latex_options = c("striped","repeat_header"),font_size=6) 
  
```

\newpage

\end{landscape}

```{r, eval = FALSE, fig.height=5, fig.width=5, fig.cap = "Species accumulation curve for all volunteers"}
#Species accumulation curve 
dat %>% group_by(collector) %>% 
  summarize(n=n(),nSpp=length(unique(genSpp))) %>%
  ggplot(aes(n,nSpp))+geom_point()+
  geom_smooth(method="lm",formula=y~sqrt(x))+
  labs(x="Number of specimens",y="Species richness")
```


```{r}
options(warn=0) #Turn warnings back on
```
