---
title: "V3_child_phenology"
author: "Rowan R"
date: "2026-02-16"
output: pdf_document
params: 
  collectorName: ""
  state: ""
  year: ""
---

```{r}

quantFun <- function(x, q) { #"Borrowed" from dplyr 1.0.0 summary
  tibble(x = quantile(x, q), q = q)
}

dataSplitEco <- data %>% st_join(ecoReg)

#Info for flight season - east
phenoE <- dataSplitEco %>%
  filter(!is.na(name)) %>%
  filter(str_detect(name,
    "Blue Mountains|Columbia Plateau|Eastern Cascades\nSlopes & Foothills|Northern Basin & Range|Snake River Plain" ))

phenoRangeE <- phenoE %>% st_drop_geometry() %>% filter(!is.na(date)) %>%
  mutate(doy=as.numeric(format(date,format="%j"))) %>%
  summarize(quantFun(x=doy,q=c(0.05,0.25,0.5,0.75,0.95))) %>%
  mutate(xOffset=as.Date(as.character(x-10),format="%j"), #x-Location for text label
         yOffset=length(unique(phenoE$genSpp[!grepl(".spp",phenoE$genSpp)])), #y-location for text label
         x=as.Date(as.character(x),format="%j"), #Actual date
         lab=paste0(q*100,"%"), # Percentile Label
         dat=format(x,format="%B %d")) #Date label

#Info for flight season - west of cascades
phenoW <- dataSplitEco %>%
  filter(!is.na(name)) %>%
  filter(str_detect(name,
                    "Cascades|Coast Range|Willamette Valley|Klamath Mountains" ))

phenoRangeW <- phenoW %>% st_drop_geometry() %>% filter(!is.na(date)) %>%
  mutate(doy=as.numeric(format(date,format="%j"))) %>%
  summarize(quantFun(x=doy,q=c(0.05,0.25,0.5,0.75,0.95))) %>%
  mutate(xOffset=as.Date(as.character(x-10),format="%j"), #x-Location for text label
         yOffset=length(unique(phenoW$genSpp[!grepl(".spp",phenoW$genSpp)])), #y-location for text label
         x=as.Date(as.character(x),format="%j"), #Actual date
         lab=paste0(q*100,"%"), # Percentile Label
         dat=format(x,format="%B %d")) #Date label

```

```{r, include=FALSE}
numAxis <- n_distinct(phenoW$genSpp)
```

## __West of__ (and including) the Cascade Mountains
Most bees (90%) were caught between `r phenoRangeW$dat[1]` and `r phenoRangeW$dat[5]`, but the peak of season (50% of specimens) was from `r phenoRangeW$dat[2]` to `r phenoRangeW$dat[4]`.

```{r sppPhenologyW, fig.cap="Phenology plot for all bee species caught in or West of the Cascade Mountains, sorted by median abundance times. Percentiles of overall emergence times (50th & 90th) are shown in grey shaded regions. Date ranges for each species (minimum, first, second, third quartiles, and maximum) are shown only for species with >10 specimens.", fig.height=11*0.6, fig.width=10*0.7}

rectDatW <- with(phenoRangeW,data.frame(perc=c("50th","90th"),
                                      ymin=x[q==0.25|q==0.05],ymax=x[q==0.75|q==0.95],
                                      xmin=c(0,0),xmax=rep(yOffset[1],2)))


phenoW %>% st_drop_geometry() %>% filter(!is.na(date)) %>%
  mutate(doy=as.numeric(format(date,format="%j"))) %>%
  filter(species!="spp.") %>% group_by(genSpp) %>%
  summarize(n=n(),quantFun(x=doy,q=seq(0,1,0.25))) %>%
  mutate(q=rank(q)) %>%
  pivot_wider(names_from=q,names_prefix="q",values_from=x) %>%
  mutate(across(c(q1,q2,q4,q5),~ifelse(n>10,.,NA))) %>%
  ungroup() %>%
  arrange(desc(q3)) %>% mutate(across(c(q1:q5),~as.Date(as.character(.),format="%j"))) %>%
  # filter(n>20) %>%
  mutate(genSpp=factor(genSpp,levels=genSpp)) %>%
  ggplot(aes(x=genSpp)) + geom_linerange(aes(ymin=q1,ymax=q5))+

  geom_linerange(aes(ymin=q2,ymax=q4),size=0.3)+ #Bee emergence dates
  geom_point(aes(y=q3),size=0.3) +
  # geom_text(data=phenoRange,aes(x=yOffset,y=xOffset,label=lab),col="red",alpha=0.8)+ #Percentile markers
  # geom_hline(data=phenoRange,aes(yintercept=x),col="red",linetype="dashed",alpha=0.8)+
  geom_rect(data=rectDatW,aes(x=NULL,xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax),alpha=0.2,show.legend = FALSE)+
  coord_flip()+labs(x=NULL,y=NULL)+
  theme(panel.grid.major.x=element_line(colour="grey"),
        #panel.grid.minor.x=element_line(colour="grey"),
        axis.text.x=element_text(size=7),
        axis.text.y=element_text(size= 1))+
  scale_y_date(date_breaks="1 month",date_labels="%B",date_minor_breaks="1 weeks")

```


\newpage

## __East of__ the Cascade Mountains
Most bees (90%) were caught between `r phenoRangeE$dat[1]` and `r phenoRangeE$dat[5]`, but the peak of season (50% of specimens) was from `r phenoRangeE$dat[2]` to `r phenoRangeE$dat[4]`.

```{r sppPhenologyE, fig.cap="Phenology plot for all bee species caught east of the Cascade Mountains, sorted by median abundance times. Percentiles of overall emergence times (50th & 90th) are shown in grey shaded regions. Date ranges for each species (minimum, first, second, third quartiles, and maximum) are shown only for species with >10 specimens.", fig.height=11*0.6, fig.width=10*0.7}

rectDatE <- with(phenoRangeE,data.frame(perc=c("50th","90th"),
                                      ymin=x[q==0.25|q==0.05],ymax=x[q==0.75|q==0.95],
                                      xmin=c(0,0),xmax=rep(yOffset[1],2)))


phenoE %>% st_drop_geometry() %>% filter(!is.na(date)) %>%
  mutate(doy=as.numeric(format(date,format="%j"))) %>%
  filter(species!="spp.") %>% group_by(genSpp) %>%
  summarize(n=n(),quantFun(x=doy,q=seq(0,1,0.25))) %>%
  mutate(q=rank(q)) %>%
  pivot_wider(names_from=q,names_prefix="q",values_from=x) %>%
  mutate(across(c(q1,q2,q4,q5),~ifelse(n>10,.,NA))) %>%
  ungroup() %>%
  arrange(desc(q3)) %>% mutate(across(c(q1:q5),~as.Date(as.character(.),format="%j"))) %>%
  # filter(n>20) %>%
  mutate(genSpp=factor(genSpp,levels=genSpp)) %>%
  ggplot(aes(x=genSpp)) + geom_linerange(aes(ymin=q1,ymax=q5))+

  geom_linerange(aes(ymin=q2,ymax=q4),size=0.3)+ #Bee emergence dates
  geom_point(aes(y=q3),size=0.3) +

  # geom_text(data=phenoRange,aes(x=yOffset,y=xOffset,label=lab),col="red",alpha=0.8)+ #Percentile markers
  # geom_hline(data=phenoRange,aes(yintercept=x),col="red",linetype="dashed",alpha=0.8)+
  geom_rect(data=rectDatE,aes(x=NULL,xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax),alpha=0.2,show.legend = FALSE)+

  coord_flip()+labs(x=NULL,y=NULL)+
  theme(panel.grid.major.x=element_line(colour="grey"),
        #panel.grid.minor.x=element_line(colour="grey"),
        axis.text.x=element_text(size=7),
        axis.text.y=element_text(size=1))+
  scale_y_date(date_breaks="1 month",date_labels="%B",date_minor_breaks="1 weeks")

```

```{r, include=FALSE}
numAxis <- n_distinct(phenoE$genSpp)
```